---
title: "First contact with R: Extract photo location"
author: Athena Hadjixenofontos, PhD
output: html_notebook
---

Note: Please use this notebook by opening the .Rmd file in RStudio, rather than working with the rendered HTML version. 

## Note on our approach

Welcome! We're taking a slightly unorthodox approach in that we'll jump right in, expose you to the look and feel of the R programming language, before explicitly defining the puzzle pieces that come together to create a working program. 

New vocabulary will be indicated by *bold* typeface. We'll be using these words a lot so you'll learn them by using them. There's no need to worry about memorizing any of their definitions at the moment. 

## Extract location from photo metadata

At this point, don't worry about being able to reproduce the code that you see here. *Instead, focus on the steps that we're taking.* Each step is a small, manageable problem that we can solve fairly easily. These small steps fit together to solve the larger problem of extracting the location where a picture was taken from the photo file. 

### Step 1: Load the photo metadata into a table that we can play with 

*Metadata* is information about information. In our example, the photos themselves are the first layer of data. On top of that we also have information such as:
* the type of camera used to take the picture
* the size of the picture measured in pixels
* the geographical coordinates of where the picture was taken (latitude and longitude)
* and lots more! 

We're interested in the latitude and longitude, since we can use those two pieces of information to pin the photo location on a map. 

The gray box below is a code block. The first and last lines tell R that everything inside this box is instructions for the computer to do things. 

Click anywhere in the code block below and press the key combination *Ctrl+Shift+Return* to run it. This executes the code. 

```{r}
# The green lines are comments. Comments help you understand what the code does. 
# Comments are not code, i.e. they don't get executed by the computer. 
# In other words, they have meaning for humans, not for computers. 
# We use them to leave little notes to our future selves or others who will use the code. 

# Load the tools that understand picture files from the "exifr" library 
library(exifr)

# Go to the folder where the pictures are saved
setwd("../photos/")

# Read in the picture metadata for all photos in the folder 
photo_metadata <- exifr(".", recursive = TRUE)
```

Notes on new vocabulary:
setwd() stands for set *working directory*. What we usually call folders on a computer can also be called directories, that's part of tha language that you'll learn as you type out commands that give the computer instructions rather than click on graphical objects to give the computer instructions. 

The parentheses indicate a *function*, and functions take inputs. For example, when you are loading the tools that can understand picture files, you have to tell the library() function the names of the tools that you want it to load! 

Some functions need more information than other functions. You see above that the exifr() function takes an *argument* called "recursive" which is passed the value "TRUE". This tells the exifr() function to look for all of the files in the working directory, and not exit after finding the first one.  

### Step 2: Find the latitude and longitude from the metadata table

Let's take a look at this table that we created. In the Environment box on the right, notice on the `photo_metadata` object under Data. 

When you click on it in a little bit, it will open up the table so you can scroll around and see what it contains. You'll have to click on the extract-photo-location.Rmd tab to get back to this notebook. 

Okay, now click on it. Whoa! There's a lot of information that comes with a picture file!

Note: The symbol NA stands for Not Available and represents missing values. Notice how we're missing a lot of information for many of the photos. That's okay, for now. 

Once you've taken a good look, execute the code block below to get a list of the names of the columns in this table. 

```{r}
names(photo_metadata)
```

We're looking for the GPSLatitude and GPSLongitude columns, since those contain the coordinates that we can use to map the location where the photo was taken. 

### Step 4: Extract only the columns with coordinates from the larger metadata table 

Exectute the code block below to extract the latitude and longitude columns, only for the photos that have the information available (i.e. excluding the photos that are missing the information). 

```{r}
# Make a new table containing only the file name, latitude and longitude columns
lat_and_long <- photo_metadata[,c("FileName", "GPSLatitude", "GPSLongitude")]

# Update the table, keeping only the rows with complete information
lat_and_long <- lat_and_long[complete.cases(lat_and_long), ]

# Let's print it out here! 
lat_and_long
```

Notes on new vocabulary:
The [ ] are called *square brackets*. They are used to get a *slice* of the larger table, You'll hear the action of "slicing" also referred to as a *subseting* operation. 

The table that we've been using is called a *data_frame*, which means that it has rows and columns and those rows and columns can have numbers or letters or words in them, i.e. mixed types of data. 

Data frames are *two-dimensional* objects, the first dimension is the rows and the second dimension is the columns. The comma inside the square brackets separates the rows and the columns, i.e. [rows, columns]. 

So when we type `photo_metadata[,c("FileName", "GPSLatitude", "GPSLongitude")]` we are saying to the computer to take all the rows (because there is nothing before the comma), and only the columns with the names FileName, GPSLatitude and GPSLongitude.  

## Woo! You did so good, getting to this point. So where were the photos taken? 

Take pairs of coordinates from 3 or 4 of the photos, and put them into Google Maps. 

To do this, open a browser window, navigate to [Google Maps](maps.google.com) and put e.g. (42.79858,	-70.808372) in the search box. 

Where was the picture taken? 

Open the photo file as well to look at the picture, does it look like it makes sense that it was taken at the location indicaded in its metadata? 




