---
title: "Where was I when I took this photo?"
author: "Athena Hadjixenofontos, PhD"
output:
  html_document: default
  html_notebook: default
---

### Welcome to R! 

In this activity, we will extract the location where a photo was taken from its metadata. 

We're taking a slightly unorthodox approach in that we'll jump right in, expose you to the look and feel of the R programming language, before explicitly defining the puzzle pieces that come together to create a working program. 

### Carry on reading, even if it doensn't all make sense initially

#### How to use this file

Please use this notebook by opening the .Rmd file in RStudio, rather than the rendered HTML version. This will allow you to run the code for yourself. 

You have all you need in these notes, if you persist you can get to the end of the activity with your sharp independence skills! 

New vocabulary will be indicated by *bold* typeface. We'll be using the new vocabulary words a lot so don't worry too much about memorizing them, you'll learn them by using them. 

#### What to focus on in this activity: problem decomposition

At this point, don't worry about being able to write for yourself the code that you see here. *Instead, focus on the steps that we're taking.* Each step is a small, manageable problem that we can solve fairly easily. These small steps fit together to solve the larger problem of extracting the location where a picture was taken from the photo file. 

### Step 1: Load the photo metadata into a table that we can play with 

*Metadata* is information about information. In our example, the photos themselves are the first layer of data. On top of that we also have information such as:
* the type of camera used to take the picture
* the size of the picture measured in pixels
* the geographical coordinates of where the picture was taken (latitude and longitude)
* and lots more! 

We're interested in the latitude and longitude, since we can use those two pieces of information to pin the photo location on a map. 

The gray box below is a code block. The first and last lines tell R that everything inside this box is instructions for the computer to do things. 

Click anywhere in the code block below and press the key combination *Ctrl+Shift+Return* to run it. This executes the code. 

```{r}
# The green lines are comments. Comments help you understand what the code does. 
# Comments are not code, i.e. they don't get executed by the computer. 
# In other words, they have meaning for humans, not for computers. 
# We use them to leave little notes to our future selves or others who will use the code. 

# Go to folder with the metadata
setwd("../datasets/")

# Load in the file that contains the photo medatata
photo_metadata <- read.table("photo-metadata.csv", header = TRUE, sep = ",")
```

#### Notes on new vocabulary:

`setwd()` stands for set *working directory*. What we usually call folders on a computer can also be called directories, that's part of tha language that you'll learn as you type out commands that give the computer instructions rather than click on graphical objects to give the computer instructions. 

The parentheses indicate a *function*, and functions take inputs. For example, when you are loading the tools that can understand picture files, you have to tell the `setwd()` function the names of the directory that you want it to go to! 

Some functions need more information than other functions. You see above that the `read.table()` function takes three *arguments*:
* the name of the file to read in
* `header` which when set to TRUE tells R that the first line of the file is not data, but the names of the columns
* `sep` which when set to "," tells R that the columns are separated by commas

### Step 2: Find the latitude and longitude from the metadata table

Let's take a look at this table that we created. In the Environment box on the right, notice on the `photo_metadata` object under Data. 

When you click on it *in a little bit*, it will open up the table so you can scroll around and see what it contains. You'll have to click on the extract-photo-location.Rmd tab to get back to this notebook. 

Okay, *now click on it*. Whoa! There's a lot of information that comes with a picture file! You can see the day and time I took these photos, where I took them and with what kind of camera. 

Note: The symbol NA stands for Not Available and represents missing values. Notice how we're missing a lot of information for many of the photos. That's okay, for now. 

Once you've taken a good look, execute the code block below to get a list of the names of the columns in this table. 

```{r}
names(photo_metadata)
```

We're looking for the GPSLatitude and GPSLongitude columns, since those contain the coordinates that we can use to map the location where the photo was taken. 

### Step 4: Extract only the columns with coordinates from the larger metadata table 

Exectute the code block below to extract the latitude and longitude columns, only for the photos that have the information available (i.e. excluding the photos that are missing the information). 

```{r}
# Make a new table containing only the file name, latitude and longitude columns
lat_and_long <- photo_metadata[,c("SourceFile", "Date", "Time", "Latitude", "Longitude")]

# Update the table, keeping only the rows with complete information
lat_and_long <- lat_and_long[complete.cases(lat_and_long), ]

# Let's print it out here! 
lat_and_long
```

#### Notes on new vocabulary:

The [ ] are called *square brackets*. They are used to get a *slice* of the larger table, You'll hear the action of "slicing" also referred to as a *subseting* operation. 

The table that we've been using is called a *data_frame*, which means that it has rows and columns and those rows and columns can have numbers or letters or words in them, i.e. mixed types of data. 

Data frames are *two-dimensional* objects, the first dimension is the rows and the second dimension is the columns. The comma inside the square brackets separates the rows and the columns, i.e. [rows, columns]. 

So when we type `photo_metadata[,c("SourceFile", "Latitude", "Longitude")]` we are saying to the computer to take all the rows (because there is nothing before the comma), and only the columns with the names FileName, Latitude and Longitude.  

## Woo-hoo! Time for the final stretch! You did so good, getting to this point. So where were the photos taken? 

Take pairs of coordinates from 3 or 4 of the photos, and put them into Google Maps. 

To do this, open a browser window, navigate to [Google Maps](maps.google.com) and put in a pair of coordinates from above inside parentheses e.g. (42.79858,	-70.808372) in the search box. 

Where was the picture taken? Where was I on that day? 

Open a Windows Explorer window, and navigate to the location of the photo. Open the photo file to look at the picture. Does it look like it makes sense that it was taken at the location indicaded by its metadata? 




