---
title: "R Notebook"
output: html_notebook
---

Welcome to your image convolution activity!

Images are divided into a grid or matrix. Each entry of the matrix is called a pixel. Each pixel has a value that corresponds with its color value. In this activity we are going to be learning how photo filters manipulate pixel values to achieve a new image. 

The library we are going to be using in this activity is Magick. Magic is a library for image processing. 

```{r}
library(magick)
```
Next, choose any image that you want to add a filter to. We are going to load the image into our workspace. You can view your image using the print function. 

```{r}
image = image_read('image.png')
print(image)
```
We are now going to apply our first filter to the image. In image processing, a kernel or convolution matrix is a small matrix. It is used for blurring, sharpening, edge detection and more. This is accomplished by doing a convolution between the image and the kernel.

Lets create our first kernel by creating a 3x3 matrix and entering the values. 
```{r}
kern1 <- matrix(0, ncol = 3, nrow = 3)
kern1[1, 2] <- 0.25
kern1[2, c(1, 3)] <- 0.25
kern1[3, 2] <- 0.25
kern1
```
Next, lets do a convolution between the image and the kernel we just created. Magick has a built in function that does the math for us.

After convolving the image, print your newly created image. What effect did this kernel have on the original image?

```{r}
img <- image_resize(image, "300x300")
img_blurred <- image_convolve(img, kern1)
image_append(c(img, img_blurred))
```
Let's create the next kernel!
```{r}
kern2 <- matrix(0, ncol = 3, nrow = 3)
kern2[1, 1] <- -2
kern2[1, 2] <- -1
kern2[2, 1] <- -1
kern2[2, c(2, 3)] <- 1
kern2[3, 2] <- 1
kern2[3, 3] <- 2
kern2
```
Convolve your image with the second kernel. What is the effect of this kernel?
```{r}
img_sharpen <- image_convolve(img, kern2)
image_append(c(img, img_sharpen))
```
Now for the final kernel ...
```{r}
kern3 <- matrix(0, ncol = 3, nrow = 3)
kern3[1, 2] <- -1
kern3[2, 2] <- 4
kern3[2, c(1, 3)] <- -1
kern3[3, 2] <- -1
kern3
```
What were your results?
```{r}
img_edge <- image_convolve(img, kern3)
image_append(c(img, img_edge))
```
Now try to experiment with creating your own kernel! What are your results?
```{r}

```

