---
title: "Is more really better?"
author: "Athena Hadjixenofontos, PhD"
date: "7/7/2017"
output: html_document
---

## Why are we doing this? 

We've been discussing the idea of bias, and to better understand it we need to calibrate our intuition by examining the factors that play into it. 

## Whether or not your observation is biased depends on the question that you ask

If you ask: "Do **all celebrities** drive Lamborghinis?", but you only look at celebrities that live in New York, you will get a biased answer that represents **"celebrities who live in New York"** more than it represents **"all celebrities"**. 

But, if you ask "Do **all celebrities who live in New York** drive Lamborghinis?" and you only look at celebrities that live in New York, you will get a more representative answer. 

A focused question is more likely to give you an accurate answer. 

It makes perfect sense, doesn't it? 

## Do more pixels in an image allow us to calculate a more accurate average color? 

In other words, does the resolution of the image influence the average color we get? 

### Setting expectations:

**If** more pixels is really better, then we will see a difference in the average colors that we calculate from the low-resolution versus high-resolution versions of the same image. 

**If** more pixels don't help us get a more accurate average color, then we won't see a differece in the average color that we calculate from the low-resolution versus high-resolution resolution versions of the same image. 

## Detective work: Setting the stage

Let's load in some images. Each image has fewer pixels than the previous version. 

We'll load in 4 images of varying resolution. Why is 4 better than 2 (a low-resolution and a high-resolution)? 

**Hint:** Think about the case where there is a difference in the average color, what would having information on the average color of 4 different photos allow us to do, that having only 2 would not? 

```{r}

```


## Detective work: The play

```{r}

```


## So, is more really better? 